CF-SuGAR Power Calculations

### 2025-03-21 Christine Chan/CF-SuGAR/power_calculations.qmd
PROC GLIMMIX for CF-SuGAR power

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Remove ages > 22 to match Lorenz et al.
annual <- annual %>% filter(Age_YrEnd < 23)

@@

- Participants over age 23 were excluded from the registry data, along with data collected prior to 2010.
- Age groups were chosen based on the Lorenz et al. data. We split the age column into equal tertiles:

```{r}
lorenz$age_group <- cut(lorenz$age, c(7, 10.48, 13.3, 22.2), right = F)
table(lorenz$age_group, lorenz$CFRD)
registry$age_group <- cut(registry$Age_YrEnd, c(7, 10.48, 13.3, 22.2), right = F)
```

# Registry base model

```{r}
registry_base <- coxph(
  Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~
    age_group + Gender,
  data = registry, cluster = eDWID
)
# Forest plot and coefficients table
registry_base %>%
  tbl_regression(exponentiate = T)
ggcoef_table(registry_base,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

The above model had a generalized $R^2$ of `r as.numeric(summary(registry_base)$rsq[1])`.

# Lorenz et al.

@@

# PROC POWER

## HR for 1 hour OGTT glucose

@@

## Regression of age group and sex on 1 hour OGTT glucose

@@

## SD of 1 hour OGTT glucose

@@

## Event probability

@@

# To Do

- PROC power has a cox reg option
  - Base model is from the registry 
  - 0.5, 1, 2 times the 1 hr glucose point estimate from Lorenz et al.
  - 80, 85, 90% power

- Mixed model power calculations for aim 3
  - Repeated measures with FEV1 and BMI
  - Mutation class and sex as predictors

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

- Age groups were chosen based on the Lorenz et al. data. We split the age column into equal tertiles:

```{r}
lorenz$age_group <- cut(lorenz$age, c(7, 10.48, 13.3, 22.2), right = F)
table(lorenz$age_group, lorenz$CFRD)
registry$age_group <- cut(registry$Age_YrEnd, c(7, 10.48, 13.3, 22.2), right = F)
# Write data
write.csv(lorenz,
  file = "./CF-SuGAR/Background/Power Calculations/Data/lorenz.csv",
  row.names = F, na = ""
)
```

# Aim 1

- Power calculations for Aim 1 used only the Lorenz et al. data and no registry data.

@@

## PROC POWER

### HR for 1 hour OGTT glucose

@@

### Regression of age group and sex on 1 hour OGTT glucose

@@

### SD of 1 hour OGTT glucose

@@

### Event probability

@@

# Aim 2

- All models below were fit using registry data.
- One BMI value of > 200 was removed

```{r}
registry$bmi[registry$bmi >= 200] <- NA
# Write for SAS
write.csv(registry,
  file = "./CF-SuGAR/Background/Power Calculations/Data/registry.csv",
  row.names = F, na = ""
)
```

## BMI

### Plot

```{r}
plot_df <- registry %>%
  select(eDWID, ReviewYear, Age_YrEnd, bmi) %>%
  drop_na()
ggplot(plot_df, aes(x = Age_YrEnd, y = bmi, group = factor(eDWID))) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme(legend.position = element_blank()) +
  theme_bw()
```

### Models

```{r}
mod1 <- lmer(bmi ~ Age_YrEnd + (1 | eDWID), data = registry)
summary(mod1,ddf="Kenward-Roger")
```

## FEV1 % predicted

### Plot

```{r}
plot_df <- registry %>%
  select(eDWID, ReviewYear, Age_YrEnd, A_FEV1_pct_predicted) %>%
  drop_na()
ggplot(plot_df, aes(
  x = Age_YrEnd, y = A_FEV1_pct_predicted, group = factor(eDWID)
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  theme(legend.position = element_blank()) +
  theme_bw()
```

### Models
